---
title: "Watershed 8 Test Case"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Before new functionality can be added to RHESSys, new code changes must be tested for their effects on the model.    
This test case is for a small watershed (W8) located in the HJ Andrews experimental forest in Oregon.

The base simulation was run for 10 years, generating basin daily output.
Use this script to run the same simulation with your new code and compare output to this base case scenario. 

1. Edit the **rhessys_ver =** to point to your new version of RHESSys.
   (and/or edit worldfile and flowtable input files if there are changes to the structure of those files )
2. Execute the commands to run RHESSys with your new code and compare the output to this base test case output with the results in:
waterbalance_diff    
percent_diff

Download the RHESSysIOinR package at https://github.com/RHESSys/RHESSysIOinR
or directly from Rstudio with
```
devtools::install_github("RHESSys/RHESSysIOinR")
```
Load the tidyverse and RHESSysIOinR libraries:
```
library(tidyverse)
library(RHESSysIOinR)
```
### Input files    
Edit the **rhessys_ver** = path and file name to point to your new version of RHESSys    
Unless new development changes are being made to the worldfile or flowtable, you will use the same input files.

```
rhessys_ver = "../rhessys_git/rhessys/rhessys7.2"
```
```{r echo = FALSE}
worldfile = "worldfiles/w8TC.world"
header = "worldfiles/w8TC.hdr"
flowtable = "flowtables/w8TC.flow"
tecfile = "tecfiles/tec.test"
```
#### Creates the command line
```
cmd1 = sprintf("%s -t %s -w %s -whdr %s -r %s",
                     rhessys_ver, tecfile, worldfile, header, flowtable)
cmd2 = sprintf("-pre out/test -s 0.355794 651.390265 -sv 0.355794 651.390265 -svalt 1.083102 1.193924, -gw 0.116316 0.916922")
cmd3 = sprintf("-st 1988 10 1 1 -ed 2000 10 1 1 -b -g")
command_line = paste(cmd1,cmd2,cmd3)
```

### Run RHESSys
```
system(command_line)
```
### Read in the test results and plot the water balance results
If your test results differ more than double the base results, something may be wrong.....
```
test = readin_rhessys_output("out/test")
test$bd = watbal_basin(test$bd)
test$bd$watbal[1]=0

par(mar=c(4.1,4.1,3.1,1.1))
plot(base$bd$date, base$bd$watbal, ylim=c(-0.000006,0.000006), ylab="Water Balance", xlab="Date")
points(test$bd$date, test$bd$watbal, col="red")
```

### Difference between the base and the test scenarios     
The **percent_diff_mean** object shows the difference in the mean streamflow, transpiration, evaporation, psn, and plant carbon results between the base and test scenarios.     
If there are large differences - are they reasonable given the changes to functionality with your new code?
```    
test.mean = test$bd %>% select(streamflow, psn, trans, evap, plantc) %>% summarize_if(is.numeric, mean)
percent_diff_mean = ((base.mean-test.mean)/test.mean)*100
percent_diff_mean
```

## Plots
#### Plot LAI
```
plot(base$bd$date, base$bd$lai, type="l", ylab="LAI", xlab="Date")
lines(test$bd$date, test$bd$lai, col="red")
```


